---
title: "Tidy Data Project"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load dependencies
```{r library}
library(purrr)
library(dplyr)
```

## Load training set, test set, and labels
A description of all the features can be found in UCI HAR Dataset/features_info.txt
Information regarding how the data was collected can be found in UCI HAR Dataset/README.txt

The resulting dataframes are train_set and test_set. These dataframe contain the data, labels, and subjects, and they also use the feature names as the names for the columns.

```{r load data}
feature_set <- read.table('UCI HAR Dataset/features.txt')
activity_labels <- read.table('UCI HAR Dataset/activity_labels.txt')

train_subjects <- read.table('UCI HAR Dataset/train/subject_train.txt')
train_X <- read.table('UCI HAR Dataset/train/X_train.txt')
train_Y <- read.table('UCI HAR Dataset/train/y_train.txt')
names(train_subjects) <- "subjectID"
names(train_X) <- feature_set$V2
names(train_Y) <- "activityLabel"
train_set <- cbind(train_X, train_subjects, train_Y)

test_subjects <- read.table('UCI HAR Dataset/test/subject_test.txt')
test_X <- read.table('UCI HAR Dataset/test/X_test.txt')
test_Y <- read.table('UCI HAR Dataset/test/y_test.txt')
names(test_subjects) <- "subjectID"
names(test_X) <- feature_set$V2
names(test_Y) <- "activityLabel"
test_set <- cbind(test_X, test_subjects, test_Y)
```

## Merge train_set and test_set
The training set and test set are merged into a single dataframe (data).
It also removes unnecessary features, so that only feature means and standard deviations remain
```{r merge}
data <- rbind(train_set, test_set)
data <- data[, grepl('std[()-]',feature_set$V2) | grepl('mean[()-]',feature_set$V2)]
for(i in 1:6){
  data[data$activityLabel == i, 'activityLabel'] <- as.character(activity_labels[activity_labels$V1 == i, 'V2'])
}
```


## Create an independent tidy data set with the average of each variable for each activity and each subject.
Using the dplyr package, a new tidy dataframe (tidy_data) with the means of each feature grouped by activity and participant. This dataframe is saved to tidy_data.txt.

```{r tidy data}
tidy_data <- data %>%
  group_by(activityLabel, subjectID) %>%
  summarize_all(mean)

write.table(tidy_data, 'tidy_data.txt', row.names = FALSE)
```